\documentclass[thesis]{thesis}
\input{preamble.tex}

\title{Distributed Complex Event Recognition}
\setthesistype{Master}
\author{Arnau Abella}
\degree{%
  Master in Innovation and Research in Informatics\\
  Advanced Computing
}
\supervisor{%
  \\
  Sergi Nadal, Universitat Politècnica de Catalunya\\
  Stijn Vansummeren, UHasselt – Hasselt University
}

\date{\today}

\begin{document}

\maketitle

\license

\acknowledgements{%
  I have so many people to thank.
  First, and foremost, I am grateful to my advisor Sergi Nadal for his enthusiasm and guidance. This thesis would have not been possible without him. Second, I am due to my co-advisor Stijn Vansummeren. His insightful remarks and suggestions have made some of the contributions of this thesis possible.
  Thanks to Marco Bucchi, Alejandro Grez, Andr\'es Quintana, and Cristian Riveros, among many others, for their numerous contributions to the field of complex event recognition that made this work possible.
  I owe a debt of gratitude to my colleague and friend, Juan Pablo Royo. He is the reason I started this master's degree after three years away from the academia.
  I cannot express enough gratitude toward my family, old and new, who have supported me in every possible way during these two harsh years. To my parents. They have supported me during the span of this master's degree. They are both remarkable parents and I am very lucky to be their son. To my brother Adri\`a. He is the best brother someone could ask for.
  I am left to thank my partner Marta. She has been with me every step on the way. No words can truly express how I feel: I love you and thank you.
}

\abstract{%
  Complex Event Recognition (CER) has emerged as a prominent technology for detecting situations of interest, in the form of query patterns, over large streams of data in real-time. Thus, having query evaluation mechanisms that minimize latency is a shared desiderata. Nonetheless, the evaluation of CER queries is well known to be computationally expensive. Indeed, such evaluation requires the maintenance of a set of partial matches which grows super-linearly in the number of processed events. While most prominent solutions for CER run in a centralized setting, this has proved inefficient for Big Data requirements, where it is necessary to scale the system to cope with an increasing arrival rate of events while maintaining the throughput. To overcome these issues, we propose a novel distributed CER system that focuses on the efficient evaluation of a large class of complex event queries, including $n$-ary predicates, time windows, and partition-by event correlation operator. This system uses a state-of-the-art automaton-based distributed algorithm that circumvents the super-linear partial match problem. Moreover, in the presence of heavy workloads, the system can scale-out by increasing the number of processing units with little overhead. We additionally provide a proof of correctness of the algorithm. We experimentally compare our system against the state-of-the-art sequential CER engine that inspired our work and show that our system outperform its predecessor in the presence of queries with complex predicates. Furthermore, we show that, in the presence of Big Data requirements, our system performance is overall better.
}

\rhead{\thepage}
\pagestyle{fancy}
\tableofcontents
\listoffigures
% \listoftables
% \listofalgorithms
% \addcontentsline{toc}{chapter}{List of Algorithm}
% \listoflistings
% \addcontentsline{toc}{chapter}{List of Source Code}
\mainmatter

\input{chapters/chapter_1.tex}
\input{chapters/chapter_2.tex}
\input{chapters/chapter_3.tex}
\input{chapters/chapter_4.tex}
\input{chapters/chapter_5.tex}
\input{chapters/chapter_6.tex}
\input{chapters/chapter_7.tex}
\input{chapters/chapter_8.tex}

\begin{appendices}
\input{chapters/appendix_A.tex}
\end{appendices}

\bibliographystyle{unsrtnat}
\bibliography{bibliography}

\end{document}
